### 2012-07-25 Doc/VRIPHYS2012/vriphys2012.tex
more on article
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@431 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{abstract} % TODO
   The ABSTRACT is to be in fully-justified italicized text, 
   between two horizontal lines,
   in one-column format, 
   below the author and affiliation information. 
   Use the word ``Abstract'' as the title, in 9-point Times, boldface type, 
   left-aligned to the text, initially capitalized. 
   The abstract is to be in 9-point, single-spaced type.
   The abstract may be up to 3 inches (7.62 cm) long. \\
   Leave one blank line after the abstract, 
   then add the subject categories according to the ACM Classification Index 
   (see http://www.acm.org/class/1998/).

% TODO
\begin{classification} % according to http://www.acm.org/class/1998/
\CCScat{Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

@@

Methods based on continuum mechanics and namely theory of elasticity gained on popularity in computer graphics and real-time simulators ever since Terzopoulos \cite{Terzopoulos1987} presented his work on elastic deformation modeling.  %% Do we
While the methods using finite elements and theory of elasticity are fairly popular in volumetric modeling \footnote{TODO: do we need ref. here?  yes but for medical applications} of deformations, they are still frowned upon in the area of thin structures. 
Yet, a special field of continuum mechanics has studied the deformations of these objects, and has proposed models based on shell theory.
This theory was integrated numerically \CD{Todo citation of a shell book} using Finite Elements, but there are two issues for obtaining interactive frame rates: 
(i) the continuum equations requires second spatial derivations that can not be handled with linear interpolation and 
(ii) the final system of equations that describes the mechanics is non-linear (if the deformations are not small).

@@

\section{Element kinematics} % {{{

\CD{intro: we propose an interpolation with a high degree of derivability while maintaining a reasonably low number of degrees of Freedom.
We introduce a new way of building the interpolation by using two stages: 
(i) a B\'ezier interpolation is defined with cubic polynomial functions 
(ii) a kinematic relation between nodes (6DOFs) defined at the vertices and the control points of the B\'ezier Triangle}

@@

B\'ezier triangles were designed in sixties by French mathematician de Casteljau.  % SURE ? What about Bezier himself ?
As with other B\'ezier or B-spline surfaces the B\'ezier triangle is
defined by a mesh of control points that do not necessarily lie on the
surface. The general $n$-th order B\'ezier triangle requires $(n + 1)(n +
2)/2$ control points and it's surface is defined as:

@@

\subsection{Kinematic between Nodes and Control Points } % {{{

The B\'ezier triangles are defined by a mesh of control points, however the
simulated object is usually described by a triangular mesh with 6 degrees of freedom per node (position \& rotation).
%
\CD{I would not say that it is "usual"... => I would say that 10 points x 3 Dofs  = 30 Dofs for a bezier triangle $\rightarrow$ it's quite a lot ! 
We propose to reduce the degrees of freedom by using 3 nodes with 6 dofs = 18 Dofs }
%
We constructthe initial B\'ezier mesh by taking into account the curvature of the triangular
mesh. Note that special care has to be taken to maintain
the continuity across the nodes and edges. During the initialization phase
we partially employ the method described in \cite{Ubach2010} to maintain
$C^0$ continuity on the edges. Each of the control points on the edge is
computed as the intersection of (\CD{Need to do a figure of this}):

@@

We do not repeat this procedure during the simulation. 
\CD{Why ?... do we argue that it is only for having fast update ?  } 
For fast update of
the control points we rigidly attach each edge point to nearest corner point (see Figure
\ref{fig-segments} for the correspondence between edge points and
associated corner points).

@@

In the initialization phase we remember the position of edge point relative to the position and orientation of the corner node:

\begin{equation}
    \overline{P_{i-j}} = (R_i^0)^T P_{i-j}^0
\end{equation}

\noindent
where $P_{i-j}^0 = P_j^0 - P_i^0$, with $j \in [4;10]$ being index of the edge node and $i \in [1;3]$ the index of associated corner node, is the segment at the beginning of the simulation and $R_i^0$ the orientation of node $P_i$.

In each step of the simulation we recompute the new position based on 
current orientation of the corner node:
\CD{Other possible explanation: we apply the rigid transformation of frame $i$ on control point $j$ }

\begin{equation}
    P_j = P_i + P_{i-j} = P_i + R_i \overline{P_{i-j}}
\end{equation}

\CD{To be consitent with what I write in following, we could introduce a notation for the dofs of the nodes:  $\mathbf{x}_i$ for the position and $\theta_i$ for the rotation of the node $i$
Then, the previous equation would be:
\begin{equation}
    P_j = \mathbf{x}_i + P_{i-j} = P_i + R(\theta_i) \overline{P_{i-j}}
\end{equation}
where $ R(\theta_i)$ is the rotation matrix for the angular position of node $i$...
}

\noindent
where $P_i$ is the current position of corner node and $R_i$ its orientation.

$C^1$ continuity on the edges can also be maintained if special care is taken
when computing the position of the central point (see \cite{Ubach2010}). 
For simplicity we compute the central point like this: 
\CD{Here it is not clear that with this formulation we do not maintain $C^1$ Continuity...}

That way if the element is not flat the central point is slightly elevated
and not in the plane of other nodes thus keeping the curvature of the
element.

>>>>

\begin{abstract} % TODO
    TODO: Abstract.

\begin{classification} % according to http://www.acm.org/class/1998/
    \CCScat{Computer Graphics}{I.3.5}{Computational Geometry and Object Modeling}{Physically based modelling};
    \CCScat{Computer Graphics}{I.3.7}{Three-Dimensional Graphics and Realism}{Animation}.
    %\CCScat{Simulation and Modeling}{I.6.8}{Types of Simulation}{Animation}
    %\CCScat{Life and Medical Sciences}{J.3}{Health}{}
\end{classification}

@@

Methods based on continuum mechanics and namely theory of elasticity gained on popularity in computer graphics and real-time simulators ever since Terzopoulos \cite{Terzopoulos1987} presented his work on elastic deformation modeling.  %% Do we
While the methods using finite elements and theory of elasticity are fairly popular in volumetric modeling of deformations \cite{Cotin1999,Nesme2006,Miller2007}, they are still frowned upon in the area of thin structures. 
Yet, a special field of continuum mechanics has studied the deformations of these objects, and has proposed models based on shell theory.
This theory was integrated numerically \CD{Todo citation of a shell book} using Finite Elements, but there are two issues for obtaining interactive frame rates: 
(i) the continuum equations requires second spatial derivations that can not be handled with linear interpolation and 
(ii) the final system of equations that describes the mechanics is non-linear (if the deformations are not small).

@@

\section{Element Kinematics} % {{{

In this subsection we propose an element featuring interpolation with high degree of flexibility while still maintaining low number of degrees of freedom.
We introduce a new way of building the interpolation by using two stages:
\begin{enumerate}
    \item A B\'ezier interpolation with cubic polynomial functions is defined on a triangle.
    \item A kinematic relation between nodes (6DOFs) defined at the vertices and the control points of the B\'ezier triangle.
\end{enumerate}

@@

B\'ezier triangles were first designed in early sixties by French mathematician de Casteljau.
As with other B\'ezier or B-spline surfaces the B\'ezier triangle is
defined by a mesh of control points that do not necessarily lie on the
surface. The general $n$-th order B\'ezier triangle requires $(n + 1)(n +
2)/2$ control points and it's surface is defined as:

@@

\subsection{Initial Construction of Control Points} % {{{

The cubic B\'ezier triangles are defined by a mesh of 10 control points each, however the
simulated object is usually described by a triangular mesh possibly with normals.
A procedure for construction of initial Bézier mesh is necessary.
We construct the initial B\'ezier mesh by taking into account the curvature of the triangular
mesh. Note that special care has to be taken to maintain
the continuity between elements across the nodes and edges, 
we partially employ the method described in \cite{Ubach2010} to maintain
$C^0$ continuity. Each of the control points on the edge is
computed as the intersection of (\CD{Need to do a figure of this}):

@@

$C^1$ continuity between elements on the edges can also be maintained if special care is taken
when computing the position of the central point
(see \cite{Ubach2010} or refer to the work of Farin \cite{Farin2002} for continuity conditions). 
For simplicity we decided to keep the edge boundary only $C^0$ continuous and we compute the central point as a function of other 9 points:

\noindent
where the points $P_i$ are named as in Figure \ref{fig-bezier}.

That way if the element is not flat the central point is slightly elevated
and not in the plane of other points thus keeping the curvature of the
element.

\subsection{Kinematic Between Nodes and Control Points } % {{{

With 10 control points and 3 DOFs per point it is a total of 30 DOFs for triangle.
This puts ... blah blah.
We propose a reduction of the number of degrees of freedom by using 3 nodes with 6 degrees of freedom
by creating a kinematic link between the control points.
Each of the edge control points is attached to nearest corner node. See Figure
\ref{fig-segments} for the correspondence between edge points and
associated corner nodes. 
In the following text the index $j \in [4;10]$ refers to edge points and index $i \in [1;3]$ to the associated corner point/node.

@@

In the initialization phase we remember the position of edge point $j$ in the frame of corner point $i$:

\begin{equation}
    \overline{P_{i-j}} = R(\theta_i^0)^T P_{i-j}^0
\end{equation}

\noindent
where
$\mathbf{x}_i$ and $\theta_i$ are the position and angular position of the node $i$,
$R(\theta_i^0)$ is the rotation matrix for the angular position of node $i$ at the beginning of the simulation and
$P_{i-j}^0 = P_j^0 - \mathbf{x}_i^0 = P_j^0 - P_i^0$ the attachment segment.

In each step of the simulation we
%recompute the new position based on current orientation of the corner node:
apply the rigid transformation of frame $i$ on control point $j$:

\begin{equation}
    P_j = \mathbf{x}_i + P_{i-j} = P_i + R(\theta_i) \overline{P_{i-j}}.
\end{equation}

For the central point we again apply interpolation of \eqref{eq-central}.
